Approximation as Differentiation notes

### 2025-02-22 notes/definitions.tex
Move LatGal to after biproducts.
Clarifying self-enrichment of CMon.

\subsection{Indexed families of objects}

For any set $I$ and any category $C$ write $\Fam(I,C)$ for the category where:
\begin{itemize}
\item objects are the $I$-indexed families of objects of $C$;
\item morphisms from $X$ to $Y$ are families of morphisms $f_i: X_i \to Y_i$ in $C$ for any $i \in I$.
\end{itemize}

\noindent Equivalently $\Fam(I,C)$ is the functor category $\Func{I}{C}$ where we interpret $I$ as a discrete
category. $\Fam(-,C): \Set^{\op} \to \Cat$ is then the functor which sends any set $I$ to $\Fam(I,C)$ and any
function $f: I \to J$ to the following \emph{reindexing functor}.

@@

The \emph{Grothendieck construction} $\Grothendieck{D}F$ for a functor $F: D \to \Cat$ is the category where:
\begin{itemize}
\item objects are pairs $(I, X)$ of an object $I$ of $D$ and an object $X$ of $FI$;
\item morphisms from $(I, X)$ to $(J, Y)$ are morphisms $f: I \to J$ in $D$ paired with a morphism $(Ff)X
\to Y$ in $FJ$.
\end{itemize}

\noindent When $F$ is contravariant, a morphism from $(I, X)$ to $(J, Y)$ is a morphism $f: I \to J$ in $D$
paired with a morphism $X \to (Ff)Y$ in $FI$. Writing $\Fam(C)$ for the Grothendieck construction
$\Grothendieck{\Set^{\op}}\Fam(-,C)$ we then have that in $\Fam(C)$:
\begin{itemize}
\item objects are pairs $(I, X)$ of a set $I$ and an indexed family $X$ in $\Fam(I,C)$;
\item morphisms from $(I, X)$ to $(J, Y)$ are functions $f: I \to J$ paired with a morphism $X \to Y[f]$ in
$\Fam(I,C)$.
\end{itemize}

@@

A commutative monoid homomorphism from $A$ to $B$ is any function $f: A \to B$ which preserves $\varepsilon$
and $\bullet$.

@@

We note that if a category $C$ is \emph{$\CMon$-enriched}, then:
\begin{enumerate}
\item every hom-object $\Hom{C}{A}{B}$ is a commutative monoid (writing $\zero_{A,B}$ for the unit and
$+_{A,B}$ for the binary operation);
\item composition is \emph{biadditive}, i.e.~given by a family of morphisms $\Hom{C}{B}{C} \tensor
\Hom{C}{A}{B} \to \Hom{C}{A}{C}$ in $\CMon$ that preserve the additive structure in $\Hom{C}{B}{C}$ and
$\Hom{C}{A}{B}$ separately:

@@

$\CMon$ is closed monoidal, and so enriched over itself; every hom-object $\CMon(A,B)$ is a commutative
monoid with:
\begin{enumerate}
\item unit given by the constant homomorphism sending every element of $A$ to $\varepsilon_B$;
\item binary operation given pointwise so that $(f + g)(a) = f(a) + f(b)$.
\end{enumerate}

@@

\begin{proposition}
If $C$ is $\CMon$-enriched then any functor category $\Func{D}{C}$ is $\CMon$-enriched.
\end{proposition}

>>>>

\begin{definition}[$\Fam(I,C)$]
For any set $I$ and any category $C$ write $\Fam(I,C)$ for the category where:
\begin{itemize}
\item objects are the $I$-indexed families of objects of $C$;
\item morphisms from $X$ to $Y$ are families of morphisms $f_i: X_i \to Y_i$ in $C$ for any $i \in I$.
\end{itemize}
\end{definition}

\noindent Equivalently $\Fam(I,C)$ is the functor category $\Func{I}{C}$ where we interpret $I$ as a discrete
category.

@@

\begin{definition}[$\Fam(-,C)$ functor]
$\Fam(-,C): \Set^{\op} \to \Cat$ is then the functor which sends any set $I$ to $\Fam(I,C)$ and any function
$f: I \to J$ to the functor $\reindex{-}{f}: \Fam(J,C) \to \Fam(I,C)$.
\end{definition}

\begin{definition}[Grothendieck construction for $F$]
Suppose a functor $F: D \to \Cat$. The \emph{Grothendieck construction} $\Grothendieck{D}F$ for $F$ is the
category where:
\begin{itemize}
\item objects are pairs $(I, X)$ of an object $I$ of $D$ and an object $X$ of $F(I)$;
\item morphisms from $(I, X)$ to $(J, Y)$ are morphisms $f: I \to J$ in $D$ paired with a morphism $F(f)(X)
\to Y$ in $F(J)$.
\end{itemize}
\end{definition}

\noindent When $F$ is contravariant, a morphism from $(I, X)$ to $(J, Y)$ is a morphism $f: I \to J$ in $D$
paired with a morphism $X \to F(f)(Y)$ in $F(I)$.

\begin{definition}[Category of families]
For any category $C$ define $\Fam(C)$ to be the Grothendieck construction
$\Grothendieck{\Set^{\op}}\Fam(-,C)$.
\end{definition}

In $\Fam(C)$:
\begin{itemize}
\item objects are pairs $(I, X)$ of a set $I$ and an indexed family $X$ in $\Fam(I,C)$;
\item morphisms from $(I, X)$ to $(J, Y)$ are functions $f: I \to J$ paired with a morphism $X \to Y[f]$ in
$\Fam(I,C)$.
\end{itemize}

@@

A commutative monoid homomorphism from $A$ to $B$ is any function $f: A \to B$ preserving $\varepsilon$ and
$\bullet$.

@@

We note that if a category $C$ is \emph{$\CMon$-enriched}, we have that:
\begin{enumerate}
\item Every hom-object $\Hom{C}{A}{B}$ is a commutative monoid; the unit $\zero_{A,B}$ is a zero morphism and
the binary operation $+_{A,B}$ is addition of morphisms.
\item Composition is \emph{biadditive}, i.e.~given by a family of morphisms $\Hom{C}{B}{C} \tensor
\Hom{C}{A}{B} \to \Hom{C}{A}{C}$ in $\CMon$ that preserve the additive structure in $\Hom{C}{B}{C}$ and
$\Hom{C}{A}{B}$ separately:

@@

$\CMon$ is closed monoidal, and so enriched over itself. Thus every hom-object $\CMon(A,B)$ is a commutative
monoid where:
\begin{enumerate}
\item unit $0$ is the constant homomorphism sending every element of $A$ to $\varepsilon_B$;
\item addition of morphisms $+$ is defined pointwise so that $+_{A}$ so that $(f + g)(a) = f(a) + g(a)$.
\end{enumerate}

@@

\begin{proposition}
If $C$ is $\CMon$-enriched then for any category $D$ the functor category $\Func{D}{C}$ is $\CMon$-enriched.
\end{proposition}
